---
title: "PUFFIN processing"
---

**Dataset preparation:**

I will add the known upstream and downstream sequences. Than includes the primers and LoxP, and part of the CDS for dsRed/EGFP. As 325pb are needed in each side, and to be independent from the genomic context of the insertion, I will complete with random sequences. Every promoter will have the same 5 random sequences upstream the LoxP.

```{r}
n_random_seq=5
up="gtacatcaagtgtatcactagcgtttaaacttaagcttccatggattacaaggatgacgatgacaagggggtacctgccccaaaaaaaaaacgcaaagtggaggacccagtaccaggatctagaggtaggtgatcctcctgctgctttggttcagggttttgcttgaggggggggggtggtgatttccttgccatgggcagactgagcagaaaaggccattgggaccatgttctgaatgcctccacctcaaccaccggccggtaggaccaaagccaccccgtgttttctcaggatctcttttcccagggagatccctcggcccaa"
down="ATAACTTCGTATAGCATACATTATACGAAGTTATATGGATCCATACTAGgacattgattattg" %>% tolower()
random=sample(c("A", "C","G","T"),size=325-nchar(down), replace=T) %>% paste0(collapse="")
fasta=readLines("Library_data/data/promoters_wo_dupl.fa")
ids=fasta[grep("^>", fasta)]
seqs=fasta[-grep("^>", fasta)]
for(i in 1:n_random_seq){
  random=sample(c("A", "C","G","T"),size=325-nchar(down), replace=T) %>% paste0(collapse="")
  for(j in 1:length(seqs)){
    id_=paste0(ids[j],"_random",i)
    id_ %>% cat(file="puffin/library_puffin.fa", append=T, sep="\n")
    paste0(random, down, seqs[j], up) %>% cat(file="puffin/library_puffin.fa", append=T,sep="\n")
  }
}
```

**Analysis**

The script puffin.py was modified to write all the outputs in folder pred

```{bash}
cd puffin
mkdir pred
mkdir pred/processed
python puffin.py sequence puffin/library_puffin.fa
```

**Processing.**

The "replicates" with different upstream random sequences will be summarized by the mean for every variable in every position. This new dataframe will be written in the folder processed.

I also calculate some statistics for each promoter:

-   The prediction score in the position=16 (the TSS as set by EPD)

-   The mean prediction score in the window 6 to 26, (-10 to 10 from TSS)

-   The relative position and value of the maximum prediction score in the sequence

As there are too many files (\~24000\*5), they are processed in chunks of FILE_BINS ids at a time and in parallel.

```{r}
process_sample=function(id){
  if(!is.na(id)){
  id=str_replace(id,"\\+", "_") %>% str_replace(">", "_")
  files=paste0("puffin/pred/", id, "_random", 1:5, ".csv")
  #files=list.files("puffin/pred", pattern=id, full.names = T) 
  obj=map(files, ~read_csv(.x, show_col_types = FALSE)%>% t() %>% as_tibble()) %>% keep(~nrow(.x)>0) 
  if(length(obj)!=5){print(paste("WARNING:", id, "with only", length(obj), "elements"))}
  for (i in 1:length(obj)){
    names(obj[[i]])=obj[[i]][1,]
    obj[[i]]=obj[[i]][-1,] %>% mutate(across(-Sequence, as.numeric))
  }
  obj=obj %>%list_rbind() %>%  group_by(Coordinate) %>% summarise(across(where(is.numeric), median), sd_prediction=sd(Prediction),Sequence=unique(Sequence), rep_random=length(obj))  %>% ungroup() %>% mutate(seq_id=id) 
  write_csv(obj, paste0("puffin/pred/processed/", id, ".csv"))
  return(obj)}
}


df_pred=data.frame()
FILE_BINS=200
plan(multisession, workers=10)
for(i in 0:(length(ids)/FILE_BINS)){ # como son demasiados archivos para abrir juntos, los separo en paquetes de FILE_BINS ids (multiplicado por n_random_seq para la cantidad de archivos)
  print(c("starting ID", i*FILE_BINS))
  m=future_map(ids[(i*FILE_BINS+1):((i+1)*FILE_BINS)] %>% str_remove("^>"),process_sample) #cargo, ordeno la data, resumo las "replicas" y guardo archivo
  g1=m %>% list_rbind()%>% group_by(seq_id) %>% filter(Prediction==max(Prediction)) %>% rename(max_score_pred_puffin=Prediction, pos_max_score_pred_puffin=Coordinate) %>% select(seq_id, max_score_pred_puffin, pos_max_score_pred_puffin)
  g2=m %>% list_rbind() %>% filter(Coordinate==252-16)%>% rename(teoTSS_score_pred_puffin=Prediction, teoTSS_inr_eff=`Sum of initiator effect`, teoTSS_motif_eff=`Sum of motif effect`, teoTSS_trinucl_eff=`Sum of trinucleotide effect`) %>% select(teoTSS_score_pred_puffin,teoTSS_inr_eff,teoTSS_motif_eff, teoTSS_trinucl_eff, seq_id)
  g3=m %>% list_rbind() %>% filter(Coordinate>252-26,Coordinate<252-6) %>%  group_by(seq_id) %>% summarise(TSS20_mean_score_pred_puffin=mean(Prediction))
  df_pred=rbind(df_pred, full_join(g2,g1) %>% full_join(g3))
} 
write_tsv(df_pred, "Analysis/Tables/Dudnyk_puffin_prediction_summ.tsv")
prom_df=left_join(prom_df, df_pred, by="seq_id")
```
